```{r}
library(broom)
library(multcomp)
library(tidyverse)
library(sandwich)
library(performance)
library(NHSRdatasets)
dt <- read.csv("poisson_sim.csv") # simulated data; number of adverse effects for three drug indications
data("LOS_model")
poissonlshtm <- readxl::read_xlsx("poissonlshtm.xlsx")
```

Example of a Poisson process

Suppose the probability that a drug produces a certain side effect 
is p = 0.1% and n = 1,000 patients in a clinical trial receive the drug. 
What is the probability 0 people experience the side effect?
  
```{r}
# The expected value is np
1000 * .001
# The probability of measuring 0 when the expected value is 1
dpois(x = 0, lambda = 1000 * .001)
options(scipen = 999, digits = 2) # sig digits

x <- 0:10
density <- dpois(x = x, lambda = 1000 * .001)
prob <- ppois(q = x, lambda = 1000 * .001, lower.tail = TRUE)
df <- data.frame(x, density, prob)

ggplot(df, aes(x = x, y = density)) +
  geom_col() +
  geom_text(
    aes(label = round(density, 2), y = density + 0.01),
    position = position_dodge(0.9),
    size = 3,
    vjust = 0
  ) +
  labs(title = "Poisson(1)",
       subtitle = "PMF and CDF of Poisson(1) distribution.",
       x = "Events (x)",
       y = "Density") +
  geom_line(data = df, aes(x = x, y = prob))

```


---------------
Poisson regression
---------------

Poisson Regression Assumptions
```{r}
# explore whether the response is a Poisson process
dt %>%
  ggplot(aes(n.ev)) +
  geom_bar() +
  facet_grid(cols = vars(ind))

# inspect means and variances
dt %>%
  group_by(ind) %>%
  summarise(mean = mean(n.ev, na.rm = T), variance = var(n.ev))

# linearity of the log(estimated mean)
mdl1 <- glm(n.ev ~ age, family = "poisson", data = dt)
plot(mdl1, 1)
fortify(mdl1) %>% 
  ggplot(aes(age, .resid)) + geom_smooth() + geom_jitter()
```

Model summary
```{r}
# obtain model summary
summary(mdl1)

# Dispersion parameter rule of thumb less than 1
sum(residuals(mdl1, type ="pearson")^2)/mdl1$df.residual

# obtain model summary
summary(mdl1)
# obtain a table
gtsummary::tbl_regression(mdl1)
```

GOF
```{r}
# Run a chi-squared test to check goodness-of-fit
pchisq(mdl1$deviance, mdl1$df.residual, lower.tail = FALSE)

# Suppose we want to drop one independent variable and compare model fit
mdl2 <- update(mdl1, . ~ . -age)
anova(mdl2, mdl1, test = "Chisq")
```

Graphical presentation of the model prediction
```{r}
# Calculate expected counts
dt$phat <- predict(mdl1, type = "response")
dt <- dt[with(dt, order(age)), ]
# and represent them graphically
ggplot(dt, aes(x = age, y = phat)) +
  geom_point(aes(y = n.ev), alpha = .5, position = position_jitter(h = .2)) +
  geom_line(size = 1) +
  labs(x = "Age", y = "Expected number of events")

```


---------------
Quasi-poisson
--------------

Overdispersion for a variable is present use quasi-Poisson:
there is more variation in the response than the model implies
We can estimate a dispersion parameter by dividing the model deviance by its corresponding degrees of freedom
or use the formal calculation dividing the Pearson residuals squared to the the residual degrees of freedom

```{r}
mdllos <- glm(LOS ~ Age + Death + Organisation, data = LOS_model, family = "poisson")
summary(mdllos)

sum(residuals(mdllos, type ="pearson")^2)/mdllos$df.residual

quasiLOS <- glm(LOS ~ Age + Death + Organisation, data = LOS_model, family = "quasipoisson")
summary(quasiLOS)
```


----------------
Negative binomial regression
----------------

An exlpicit likelihood alternative to quasi-poisson in the case of overdispersion 
which allows for an additional parameter. As a result, it yields a more flexible model.

```{r}
library(MASS)
negbnLOS <- glm.nb(LOS ~ Age + Death + Organisation, data = LOS_model)
summary(negbnLOS)
```


Since there are multiple levels in the Organisation variable, we should adjust 
for multiple comparisons using Tukey's honestly significant difference
When there are statistically significant differences, interaction terms must be considered.

```{r}
# linfct is a specification of a linear hypothesis
# mcp(var = "method")
summary(glht(mdllos, linfct = mcp(Organisation = "Tukey")))
```



------------------
Zero-inflated models
------------------

In settings with excess zero counts, the Poisson model is fitted with an 
additional parameter which corresponds to the proportion of zeroes.


```{r}
library(pscl)
poisalc <- glm(death ~ agein + alcohol + drug.use, data = poissonlshtm, family = "poisson")
summary(poisalc)
zimalc <- zeroinfl(death ~ agein + alcohol + drug.use, data = poissonlshtm)
check_zeroinflation(poisalc, tolerance = 0.05)
summary(zimalc)
# Vuong test
vuong(zimalc, poisalc)
```
